Debugging by \textit{lastChange}

### 2010-10-15 papers/pldi2011/sigplanconf-template.tex

To attack the mentioned problem, we propose a new functionality in debuggers, \textit{lastChange}, which locates the last place a value is changed and lets developer to collect data and ask for more \textit{lastChange} queries at the located point. While \textit{lastChange} algorithm is based on program re-execution, contrary to other replay-based algorithms which require exactly the same re-executions, it only requires \textit{bug reproduciblity} (i.e., a test case is available which reproduces the bug).

As a proof of concept we developed \textit{RevNav}, a JavaScript prototype, which adds the \textit{lastChange} feature to Firebug JavaScript debugger. Moreover, \textit{RevNav} provides mechanism for automated bug reproduction, and a centralized user interface which summarizes investigated execution points and collected results.

% how we buil it on breakpoint technology
\end{abstract}

@@

Locating defects cause the bug is the main part of debugging. A common strategy for locating defects is starting from bug symptoms and backward movement by following the origings to wrong values. While this strategy seems straight-forward and effective, applying this strategy using traditional methods is complicated. There are two traditional approaches to debugging, breakpoint-based and log-based debugging. None of these approaches assist developer in finding origings to a wrong value. It is due to developer to search source files, find the list of possible origings to a wrong value and set breakpoints or insert log statements in a way that covers all possible origins.

The next dilemma is data collection. In breakpoint debugging, developer has to memorize values or collect data manually at every breakpoint hit and in cases that there are many breakpoint hits, this becomes a tedious task. In log-based debugging, developer has to decide about data should be collected when inserts the log statement. It is very common that the data developer decided to be collected is insufficient. Developer has to change the log statements and re-executes the program another time. Once the sufficient data is collected, it still requires analysis and understanding. Developer usually ends up in dealing with long log files and analyzing huge collected data.

@@

%---------------------------------------------------------------------------------------------------
\section{Introductory example}
We illustrate the \textit{lastChange} functionality by a simple example. The example demonstrates a buggy JavaScript code in a html page (Figure ~\ref{fig:js-code}). The page contains a button (line 40) which shows the \texttt{myObject.myProperty} value. It is supposed that whenever user clicks on the button, the value of \texttt{myObject.myProperty} and therefore button's label is increased by one. Once the page is loaded for the first time the button shows \texttt{1} as the initial value of \texttt{myObject.myProperty}. The bug happens when user clicks on the button and instead of \texttt{2}, \texttt{0} appears.

Once user clicks on the button, \texttt{onClick} function (line 13) is called. This function increases the value of \texttt{myObject.myProperty} by one (line 15) and calls \texttt{updateButton} function which updates button's text to the new value (line 22). To start debugging, the developer sets a abreakpoint on line 22 and once the button is clicked execution is paused at line 22. (Figure ~\ref{fig:example1}) shows the debugger while the execution is paused. The developer can examine the program state at the right panel and as it shows, the \texttt{myObject.myProperty} value at this point is zero which is obviously a wrong value.

To apply backward search strategy for locating defects, the developer first needs to know the origin to the wrong value. To achieve this goal using breakpoints, the developer should search code to find and set breakpoint at all possible places that \texttt{myObject.myProperty} might get a new value. Even if the developer could sucessfully finish this step (There is no simple way to check it), the next step is monitoring the re-execution and collecting data at breakpoint hits before reaching line 22. All this work is useless if an origin was missed in the first step or if the data was collected in the second step is insufficient.

>>>>

To address this problem, we propose a new functionality in debuggers, \textit{lastChange}, which locates the last place a value is changed and lets developer to collect data and ask for more \textit{lastChange} queries at the located point. Although \textit{lastChange} algorithm is based on program re-execution, contrary to other replay-based algorithms which require exactly the same re-executions, it only requires bug reproduction by re-executions. In other words \textit{bug reproduciblity} (i.e., a test case is available which reproduces the bug) is the only prerequisite for \textit{lastChange} functinality.

As a proof of concept, we developed \textit{RevNav}, a prototype which enhances Firebug JavaScript debugger with \textit{lastChange} feature. Moreover, \textit{RevNav} provides mechanisms for automated bug reproduction, and a novel user interface to investigated execution points and results.

\end{abstract}

@@

Locating defects cause the bug is the main part of debugging. A common strategy for locating defects is starting from bug symptoms and backward movement by following the origings to wrong values. While this strategy seems straightforward and effective, applying this strategy using traditional methods is complicated. There are two traditional approaches to debugging, breakpoint-based and log-based debugging. None of these approaches assist developer in finding origings to a wrong value. It is due to developer to search source files, find the list of possible origings to a wrong value and set breakpoints or insert log statements in a way that covers all possible origins.

The next dilemma is data collection. In breakpoint debugging, developer has to memorize values or manually collect data at every breakpoint hit. As the number of breakpoint hits increases, the process of checking the program state, collecting data and resuming the execution becomes cumbersome. While in breakpoint-based debugging, the whole program state is available to developer, in log-based debugging, developer has to decide about data should be collected when inserts the log statement. It is very common that developer has to repeat this step several times due to insufficient collected data. Once the adequate data is collected, it still requires analysis and understanding. Developers usually end up in dealing with long log files and analyzing huge collected data.

@@

%---------------------------------------------------------------------------------------------------
\section{Introductory example}
We illustrate the \textit{lastChange} functionality by a simple example. The example demonstrates a buggy JavaScript code in a html page (Figure ~\ref{fig:js-code}). The page contains a button (line 40) which shows the value of \texttt{myObject.myProperty}. It is supposed that whenever user clicks on the button, the value of \texttt{myObject.myProperty} and therefore button's label is increased by one. Once the page is loaded for the first time the button shows \texttt{1} as the initial value of \texttt{myObject.myProperty}. The bug happens when user clicks on the button and instead of \texttt{2}, \texttt{0} appears.

Once user clicks on the button, \texttt{onClick} function (line 13) is called. This function increases the value of \texttt{myObject.myProperty} by one (line 15) and calls \texttt{updateButton} function which updates button's text to the new value (line 22). To start debugging, the developer sets a breakpoint on line 22 and once the button is clicked execution is paused at line 22. Figure ~\ref{fig:example1} shows the debugger while the execution is paused. The developer can examine the program state at the right panel and as it shows, the \texttt{myObject.myProperty} value at this point is zero which is obviously a wrong value.

To apply backward search strategy for locating defects, the developer first needs to know the origin to the wrong value. To achieve this goal using breakpoints, the developer should search code to find and set breakpoint at all possible places that \texttt{myObject.myProperty} might get a new value. Having the fact that an object and property can be accessed and changed through different names and methods, we realize that the difficulty of this task. Even if the developer could sucessfully finish this step (There is no simple way to check it!), the next step is monitoring the re-execution and collecting data at breakpoint hits before reaching line 22. All this work should be repeated if an origin was missed in the first step or if the data was collected in the second step is insufficient.
